An equivalence of compositional frameworks

### 2018-10-23 EquivCompFrameworks_20180911kc.tex
changing bold's to mathbf's

% An equivalence of compositional frameworks
% John Baez, Kenny Courser and Christina Vasilakopoulou
% 2019/09/11 - KC
\documentclass{amsart}
\input{Equiv_preamble.tex}

@@

\begin{document}
\maketitle

\begin{center}   
 % {\bf Structured cospans \\}   
  \vspace{0.3cm}
  {\em John\ C.\ Baez \\}
  \vspace{0.3cm}
  {\small
 Department of Mathematics \\
    University of California \\
  Riverside CA, USA 92521 \\ and \\
 Centre for Quantum Technologies  \\
  National University of Singapore \\
    Singapore 117543  \\    } 
 \vspace{0.4cm}
{\em Kenny Courser \\}
\vspace{0.3cm}
   {\small
  Department of Mathematics \\
  University of California \\
  Riverside CA, USA 92521 \\}
  \vspace{0.4cm}   
{\em Christina Vasilakopoulou \\}
\vspace{0.3cm}
   {\small
  Department of Mathematics \\
  University of California \\
  Riverside CA, USA 92521 \\}
  \vspace{0.3cm}   
  {\small email: baez@math.ucr.edu, courser@math.ucr.edu, vasilak@ucr.edu\\} 
  \vspace{0.3cm}   
  {\small \today}
  \vspace{0.3cm}   
\end{center}

\begin{abstract}
\noindent
The first two authors have developed a compositional framework well-suited for studying networks that are built out of finite sets equipped with extra stuff. This framework, which goes by the name of `structured cospans', utilizes double categories where the objects are seen as inputs and outputs, horizontal 1-cells are `open networks', and 2-morphisms are maps between open networks. In this setup a functor $L \colon \textrm{A} \to \textrm{X}$, which is typically a left adjoint, is used to replace the objects and vertical 1-morphisms of a given double category $\mathbb{X}$ with the objects and morphisms, respectively, of the category $\textrm{A}$. Horizontal 1-cells are then cospans in $\textrm{X}$ of a particular form with 2-morphisms given by maps of these cospans. Fong has also developed a similar framework utilizing cospans to study open networks which goes by the name of `decorated cospans'. In this setup, a lax monoidal functor $F \colon \textrm{A} \to \textrm{Set}$ is used to `decorate' the apices of cospans in $\textrm{A}$ with elements of $\textrm{Set}$ giving the cospans extra structure. Using a slight variation of Fong's framework, we prove that these two frameworks are equivalent in the situation where a left adjoint can be obtained from a lax monoidal pseudofunctor using a well known construction of Grothendieck.
\end{abstract}

@@

One approach to remedying this nuisance is to instead view $F$ as a functor $F \colon \textrm{A} \to \bold{Cat}$ and then consider $F(c)$ not as a \emph{set} of decorations on the element $c$ but as a \emph{category} of decorations on the element $c$. Then one can construct a similar symmetric monoidal category similar to Fong's, which will be called $D(H(\mathbb{F}\textnormal{Cospan}(\mathrm{A})))$ for reasons that will be explained later. This symmetric monoidal category will have:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$ and}
\item{morphisms as isomorphism classes of cospans of $\mathrm{A}$ together with an element of image of the apex under the functor $F$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Except now, given another morphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two morphisms are in the same isomorphism class if the following diagrams commute
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (B') at (1,-1) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(c)$};
\node (E') at (4,-0.5) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \colon F(f)(d) \to d^\prime$ in the category $F(c^\prime)$.
}
\end{enumerate}
Note that the existence of an isomorphism $\iota \colon F(f)(d) \to d^\prime$ is equivalent to the triangle above on the right commuting up to a 2-isomorphism.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (4,0) {$\iota \Swarrow$};
\node (D) at (3,0) {$1$};
\node (E) at (4.5,1) {$F(c)$};
\node (E') at (4.5,-1) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
In other words, the isomorphism $\iota$ is a \emph{natural} isomorphism. We will typically just write this as $\iota \colon F(f)(d) \to d^\prime$ to conserve space.

@@

Let $\mathrm{A}$ be a category with finite colimits and $F \colon \mathrm{A} \to \bold{Cat}$ a lax monoidal pseudofunctor. Then there exists a double category $\mathbb{F} \textnormal{Cospan}(\mathrm{A})$ which has:

@@

where $!_c \in F(c)$ is the trivial decoration on $c$ given by the composition of the unique map $F(!) \colon F(0) \to F(c)$ and the morphism $\phi \colon 1 \to F(0)$  which comes from the symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \bold{Cat}$. For morphisms, the structure functor $U$ is defined as:

@@

Given two composable horizontal 1-cells:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (1,-0.5) {$d \in F(c)$};
\node (E) at (3,0) {$b$};
\node (F) at (4,0) {$c^\prime$};
\node (G) at (5,0) {$e$};
\node (H) at (4,-0.5) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(E) edge node[above]{$i^\prime$} (F)
(G) edge node[above]{$o^\prime$} (F);
\end{tikzpicture}
\]
the composite is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,1) {$c^\prime$};
\node (E) at (4,0) {$e$};
\node (F) at (2,2) {$c+c^\prime$};
\node (G) at (2,3) {$c+_{b} c^\prime$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(C) edge node [above] {$i^\prime$} (D)
(E) edge node [above] {$o^\prime$} (D)
(B) edge node [above] {$j$} (F)
(D) edge node [above] {$j^\prime$} (F)
(F) edge node [left] {$\psi$} (G)
(A) edge node [left] {$\psi j i$} (G)
(E) edge node [right] {$\psi j^\prime o^\prime$} (G);
\end{tikzpicture}
\]
with the corresponding decoration of the apex $\hat{d} \in F(c+_b c^\prime)$ given by:
$$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{d \times d^\prime} F(c) \times F(c^\prime) \xrightarrow{\phi_{c,c^\prime}} F(c+c^\prime) \xrightarrow{F(\psi)} F(c+_{b}c^\prime)$$
where $\psi \colon c + c^\prime \to c+_b c^\prime$ is the natural map from the coproduct to the pushout and $\phi_{c,c^\prime} \colon F(c) \times F(c^\prime) \to F(c+c^\prime)$ is the natural transformation of the symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \bold{Cat}$. Denoting the first and second of these horizontal 1-cells as $M$ and $N$, respectively, the source and target structure functors satisfy the equations $S(N \odot M)=S(M)$ and $T(N \odot M)=T(N)$.

@@

and then multiply these viewed as two 2-morphisms in $\bold{Cat}$ which results in:

@@

As is usual concerning the interchange law of double categories of this nature, only the `interior' of the two compositions appears different, but the two morphisms $(h_1^\prime +_{g^\prime} h_2^\prime)(h_1 +_g h_2) \colon c+_b e \to c'' +_{b''} e''$ and $(h_1^\prime h_1) +_{g^\prime g} (h_2^\prime h_2) \colon c+_b e \to c'' +_{b''}e''$ are the same universal map realized in two different ways. The two morphisms of decorations are two compositions of 2-morphisms in $\bold{Cat}$ for which the interchange law already holds, and as a result, the morphisms $$(\iota_{1,2}^\prime \iota_{1,2}) \colon F((h_1^\prime +_{g^\prime} h_2^\prime)(h_1 +_g h_2))(d_{1,2}) \to d_{1,2}''$$ and $$(\iota_1^\prime \iota_1,\iota_2^\prime \iota_2) \colon F((h_1^\prime h_1)+_{g^\prime g} (h_2^\prime h_2))(d_{1,2}) \to d_{1,2}''$$ are also the same.

@@

Let $\mathrm{A}$ be a category with finite colimits and $F \colon \mathrm{A} \to \bold{Cat}$ a symmetric lax monoidal pseudofunctor. Then the double category $\mathbb{F}\textnormal{Cospan}(\mathrm{A})$ is symmetric monoidal.

@@

where $0$ is the monoidal unit of $\mathrm{A}$ and $! \colon 1 \to F(0)$ is the morphism which is part of the structure of the symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \bold{Cat}$. Tensoring an object with the monoidal unit, say, on the left:

@@

%Again, let $F \colon \bold{FinSet} \to \bold{Set}$ be the symmetric lax monoidal functor that maps a finite set $N$ to the large set of all possible graph structures whose underlying set of vertices is $N$. Then a result of Fong \cite{Fong} yields a symmetric monoidal category $F\textnormal{Cospan}(\bold{FinSet})$ which has:
%\begin{enumerate}
%\item{finite sets as objects and}
%\item{isomorphism classes of `open' graphs as morphisms, where an open graph is a cospan of finite sets whose apex is equipped with the structure of a graph.}
%\end{enumerate}

@@

\begin{cor}
Given a symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \bold{Cat}$ where $\mathrm{A}$ is a category with finite colimits and whose monoidal structure is given by binary coproducts, there exists a symmetric monoidal category $D(H(\mathbb{F}\textnormal{Cospan}(\mathrm{A})))$ which has:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans of $\mathrm{A}$, where an $F$-decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Given another $F$-decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism $f \colon c \to c^\prime$ such that following diagram commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B') at (1,0.5) {$c$};
\node (B) at (1,-0.5) {$c^\prime$};
\node (C) at (2,0) {$b$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B)
(A) edge node[above]{$i$} (B')
(C) edge node[above]{$o$} (B')
(B') edge node[left]{$f$} (B);
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \colon F(f)(d) \to d^\prime$ in $F(c^\prime)$.}
\end{enumerate}
\end{cor}
In this symmetric monoidal category, isomorphism classes are as they should morally be, and the instance of two graphs having different edge sets does not prevent them from being in the same isomorphism class due to the isomorphism $\iota$.

@@

Let $F \colon \mathrm{A} \to \bold{Cat}$ be a symmetric lax monoidal pseudofunctor. Then by Theorem \ref{main1} of the previous section, we get a symmetric monoidal double category $\mathbb{F}\textnormal{Cospan}(\mathrm{A})$. This symmetric monoidal double category has:

@@

Given another symmetric lax monoidal pseudofunctor $F^\prime \colon \mathrm{A^\prime} \to \bold{Cat}$, we can obtain another symmetric monoidal double category $\mathbb{F^\prime}\textnormal{Cospan}(\mathrm{A^\prime})$. Then a map from $\mathbb{F}\textnormal{Cospan}(\mathrm{A})$ to $\mathbb{F^\prime}\textnormal{Cospan}(\mathrm{A^\prime})$ will be a double functor $\mathbb{H} \colon \mathbb{F}\textnormal{Cospan}(\mathrm{A}) \to \mathbb{F^\prime}\textnormal{Cospan}(\mathrm{A^\prime})$ whose object component is given by a finite colimit preserving functor $\mathbb{H}_0 = H \colon \mathrm{A} \to \mathrm{A^\prime}$ and whose arrow component is given by a functor $\mathbb{H}_1$ defined on horizontal 1-cells by:

@@

where $E \colon \bold{Cat} \to \bold{Cat}$ is a terminal category preserving 2-functor such that the following diagram commutes:

@@

\node (B) at (1,0) {$\bold{Cat}$};

@@

\node (D) at (1,-1) {$\bold{Cat}$};

@@

Recall that we can think of the element $d \in F(c)$ as a morphism $d \colon 1 \to F(c)$ and the morphism $\iota \colon F(h)(d) \to d^\prime$ of $F(c^\prime)$ as a 2-morphism  in $\bold{Cat}$:

@@

Applying the terminal category preserving 2-functor $E \colon \bold{Cat} \to \bold{Cat}$ to this diagram yields:

@@

\section{From pseudofunctors to left adjoints}
In this section we investigate the necessary conditions to obtain a category $\mathrm{X}$ and a left adjoint $L \colon \mathrm{A} \to \mathrm{X}$ from a category $\mathrm{A}$ with finite colimits and a symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \bold{Cat}$. The main tool that allows us to do this will be the Grothendieck construction which we recall:

\begin{defn}
Let $\bold{Cat}_{\textnormal{lax},\star}$ denote the \define{2-category of lax-pointed categories} which has:
\begin{enumerate}
\item{objects as pairs $(\mathrm{A},c)$ where $\mathrm{A}$ is a category and $c$ is an object of $\mathrm{A}$, and}
\item{a morphism from $(\mathrm{A},c)$ to $(\mathrm{X},y)$ is a pair $(F,f)$ where $F \colon \mathrm{A} \to \mathrm{X}$ is a functor and $f \colon F(c) \to y$ is a morphism in $\mathrm{X}$.}
\end{enumerate}
\end{defn}

>>>>

% An equivalence of compositional frameworks
% John Baez, Kenny Courser and Christina Vasilakopoulou
% 2019/09/11 - KC
\documentclass{amsart}
\input{Equiv_preamble.tex}
\usepackage[foot]{amsaddr}
\usepackage{comment}

@@

\author{John\ C.\ Baez$^{*,\bullet}$, Kenny Courser$^*$, and Christina Vasilakopoulou$^*$}
\address{$^*$Department of Mathematics, University of California, Riverside CA, USA 92521}
\address{$^\bullet$Centre for Quantum Technologies, National University of Singapore, Singapore 117543}
\email{baez@ucr.edu, kcour001@ucr.edu, vasilak@ucr.edu}

\begin{document}
\begin{abstract}
\noindent
The first two authors have developed a compositional framework well-suited for studying networks that are built out of finite sets equipped with extra stuff. This framework, which goes by the name of `structured cospans', utilizes double categories where the objects are seen as inputs and outputs, horizontal 1-cells are `open networks', and 2-morphisms are maps between open networks. In this setup a functor $L \colon \textrm{A} \to \textrm{X}$, which is typically a left adjoint, is used to replace the objects and vertical 1-morphisms of a given double category $\mathbb{X}$ with the objects and morphisms, respectively, of the category $\textrm{A}$. Horizontal 1-cells are then cospans in $\textrm{X}$ of a particular form with 2-morphisms given by maps of these cospans. Fong has also developed a similar framework utilizing cospans to study open networks which goes by the name of `decorated cospans'. In this setup, a lax monoidal functor $F \colon \textrm{A} \to \textrm{Set}$ is used to `decorate' the apices of cospans in $\textrm{A}$ with elements of $\textrm{Set}$ giving the cospans extra structure. Using a slight variation of Fong's framework, we prove that these two frameworks are equivalent in the situation where a left adjoint can be obtained from a lax monoidal pseudofunctor using a well known construction of Grothendieck.
\end{abstract}

\maketitle

\begin{comment}
\begin{center}   
 % {\bf Structured cospans \\}   
  \vspace{0.3cm}
  {\em John\ C.\ Baez \\}
  \vspace{0.3cm}
  {\small
 Department of Mathematics \\
    University of California \\
  Riverside CA, USA 92521 \\ and \\
 Centre for Quantum Technologies  \\
  National University of Singapore \\
    Singapore 117543  \\    } 
 \vspace{0.4cm}
{\em Kenny Courser \\}
\vspace{0.3cm}
   {\small
  Department of Mathematics \\
  University of California \\
  Riverside CA, USA 92521 \\}
  \vspace{0.4cm}   
{\em Christina Vasilakopoulou \\}
\vspace{0.3cm}
   {\small
  Department of Mathematics \\
  University of California \\
  Riverside CA, USA 92521 \\}
  \vspace{0.3cm}   
  {\small email: baez@math.ucr.edu, courser@math.ucr.edu, vasilak@ucr.edu\\} 
  \vspace{0.3cm}   
  {\small \today}
  \vspace{0.3cm}   
\end{center}   
\end{comment}

\tableofcontents

@@

One approach to remedying this nuisance is to instead view $F$ as a functor $F \colon \textrm{A} \to \mathbf{Cat}$ and then consider $F(c)$ not as a \emph{set} of decorations on the element $c$ but as a \emph{category} of decorations on the element $c$. Then one can construct a similar symmetric monoidal category similar to Fong's, which will be called $D(H(\mathbb{F}\textnormal{Cospan}(\mathrm{A})))$ for reasons that will be explained later. This symmetric monoidal category will have:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$ and}
\item{morphisms as isomorphism classes of cospans of $\mathrm{A}$ together with an element of image of the apex under the functor $F$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Except now, given another morphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two morphisms are in the same isomorphism class if the following diagrams commute
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (B') at (1,-1) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(c)$};
\node (E') at (4,-0.5) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \colon F(f)(d) \to d^\prime$ in the category $F(c^\prime)$.
}
\end{enumerate}
Note that the existence of an isomorphism $\iota \colon F(f)(d) \to d^\prime$ is equivalent to the triangle above on the right commuting up to a 2-isomorphism.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (4,0) {$\iota \Swarrow$};
\node (D) at (3,0) {$1$};
\node (E) at (4.5,1) {$F(c)$};
\node (E') at (4.5,-1) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
In other words, the isomorphism $\iota$ is a \emph{natural} isomorphism. We will typically just write this as $\iota \colon F(f)(d) \to d^\prime$ to conserve space.

@@

Let $\mathrm{A}$ be a category with finite colimits and $F \colon \mathrm{A} \to \mathbf{Cat}$ a lax monoidal pseudofunctor. Then there exists a double category $\mathbb{F} \textnormal{Cospan}(\mathrm{A})$ which has:

@@

where $!_c \in F(c)$ is the trivial decoration on $c$ given by the composition of the unique map $F(!) \colon F(0) \to F(c)$ and the morphism $\phi \colon 1 \to F(0)$  which comes from the symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \mathbf{Cat}$. For morphisms, the structure functor $U$ is defined as:

@@

Given two composable horizontal 1-cells:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (1,-0.5) {$d \in F(c)$};
\node (E) at (3,0) {$b$};
\node (F) at (4,0) {$c^\prime$};
\node (G) at (5,0) {$e$};
\node (H) at (4,-0.5) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(E) edge node[above]{$i^\prime$} (F)
(G) edge node[above]{$o^\prime$} (F);
\end{tikzpicture}
\]
the composite is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,1) {$c^\prime$};
\node (E) at (4,0) {$e$};
\node (F) at (2,2) {$c+c^\prime$};
\node (G) at (2,3) {$c+_{b} c^\prime$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(C) edge node [above] {$i^\prime$} (D)
(E) edge node [above] {$o^\prime$} (D)
(B) edge node [above] {$j$} (F)
(D) edge node [above] {$j^\prime$} (F)
(F) edge node [left] {$\psi$} (G)
(A) edge node [left] {$\psi j i$} (G)
(E) edge node [right] {$\psi j^\prime o^\prime$} (G);
\end{tikzpicture}
\]
with the corresponding decoration of the apex $\hat{d} \in F(c+_b c^\prime)$ given by:
$$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{d \times d^\prime} F(c) \times F(c^\prime) \xrightarrow{\phi_{c,c^\prime}} F(c+c^\prime) \xrightarrow{F(\psi)} F(c+_{b}c^\prime)$$
where $\psi \colon c + c^\prime \to c+_b c^\prime$ is the natural map from the coproduct to the pushout and $\phi_{c,c^\prime} \colon F(c) \times F(c^\prime) \to F(c+c^\prime)$ is the natural transformation of the symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \mathbf{Cat}$. Denoting the first and second of these horizontal 1-cells as $M$ and $N$, respectively, the source and target structure functors satisfy the equations $S(N \odot M)=S(M)$ and $T(N \odot M)=T(N)$.

@@

and then multiply these viewed as two 2-morphisms in $\mathbf{Cat}$ which results in:

@@

As is usual concerning the interchange law of double categories of this nature, only the `interior' of the two compositions appears different, but the two morphisms $(h_1^\prime +_{g^\prime} h_2^\prime)(h_1 +_g h_2) \colon c+_b e \to c'' +_{b''} e''$ and $(h_1^\prime h_1) +_{g^\prime g} (h_2^\prime h_2) \colon c+_b e \to c'' +_{b''}e''$ are the same universal map realized in two different ways. The two morphisms of decorations are two compositions of 2-morphisms in $\mathbf{Cat}$ for which the interchange law already holds, and as a result, the morphisms $$(\iota_{1,2}^\prime \iota_{1,2}) \colon F((h_1^\prime +_{g^\prime} h_2^\prime)(h_1 +_g h_2))(d_{1,2}) \to d_{1,2}''$$ and $$(\iota_1^\prime \iota_1,\iota_2^\prime \iota_2) \colon F((h_1^\prime h_1)+_{g^\prime g} (h_2^\prime h_2))(d_{1,2}) \to d_{1,2}''$$ are also the same.

@@

Let $\mathrm{A}$ be a category with finite colimits and $F \colon \mathrm{A} \to \mathbf{Cat}$ a symmetric lax monoidal pseudofunctor. Then the double category $\mathbb{F}\textnormal{Cospan}(\mathrm{A})$ is symmetric monoidal.

@@

where $0$ is the monoidal unit of $\mathrm{A}$ and $! \colon 1 \to F(0)$ is the morphism which is part of the structure of the symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \mathbf{Cat}$. Tensoring an object with the monoidal unit, say, on the left:

@@

%Again, let $F \colon \mathbf{FinSet} \to \mathbf{Set}$ be the symmetric lax monoidal functor that maps a finite set $N$ to the large set of all possible graph structures whose underlying set of vertices is $N$. Then a result of Fong \cite{Fong} yields a symmetric monoidal category $F\textnormal{Cospan}(\mathbf{FinSet})$ which has:
%\begin{enumerate}
%\item{finite sets as objects and}
%\item{isomorphism classes of `open' graphs as morphisms, where an open graph is a cospan of finite sets whose apex is equipped with the structure of a graph.}
%\end{enumerate}

@@

\begin{cor}
Given a symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \mathbf{Cat}$ where $\mathrm{A}$ is a category with finite colimits and whose monoidal structure is given by binary coproducts, there exists a symmetric monoidal category $D(H(\mathbb{F}\textnormal{Cospan}(\mathrm{A})))$ which has:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans of $\mathrm{A}$, where an $F$-decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Given another $F$-decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism $f \colon c \to c^\prime$ such that following diagram commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B') at (1,0.5) {$c$};
\node (B) at (1,-0.5) {$c^\prime$};
\node (C) at (2,0) {$b$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B)
(A) edge node[above]{$i$} (B')
(C) edge node[above]{$o$} (B')
(B') edge node[left]{$f$} (B);
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \colon F(f)(d) \to d^\prime$ in $F(c^\prime)$.}
\end{enumerate}
\end{cor}
In this symmetric monoidal category, isomorphism classes are as they should morally be, and the instance of two graphs having different edge sets does not prevent them from being in the same isomorphism class due to the isomorphism $\iota$.

@@

Let $F \colon \mathrm{A} \to \mathbf{Cat}$ be a symmetric lax monoidal pseudofunctor. Then by Theorem \ref{main1} of the previous section, we get a symmetric monoidal double category $\mathbb{F}\textnormal{Cospan}(\mathrm{A})$. This symmetric monoidal double category has:

@@

Given another symmetric lax monoidal pseudofunctor $F^\prime \colon \mathrm{A^\prime} \to \mathbf{Cat}$, we can obtain another symmetric monoidal double category $\mathbb{F^\prime}\textnormal{Cospan}(\mathrm{A^\prime})$. Then a map from $\mathbb{F}\textnormal{Cospan}(\mathrm{A})$ to $\mathbb{F^\prime}\textnormal{Cospan}(\mathrm{A^\prime})$ will be a double functor $\mathbb{H} \colon \mathbb{F}\textnormal{Cospan}(\mathrm{A}) \to \mathbb{F^\prime}\textnormal{Cospan}(\mathrm{A^\prime})$ whose object component is given by a finite colimit preserving functor $\mathbb{H}_0 = H \colon \mathrm{A} \to \mathrm{A^\prime}$ and whose arrow component is given by a functor $\mathbb{H}_1$ defined on horizontal 1-cells by:

@@

where $E \colon \mathbf{Cat} \to \mathbf{Cat}$ is a terminal category preserving 2-functor such that the following diagram commutes:

@@

\node (B) at (1,0) {$\mathbf{Cat}$};

@@

\node (D) at (1,-1) {$\mathbf{Cat}$};

@@

Recall that we can think of the element $d \in F(c)$ as a morphism $d \colon 1 \to F(c)$ and the morphism $\iota \colon F(h)(d) \to d^\prime$ of $F(c^\prime)$ as a 2-morphism  in $\mathbf{Cat}$:

@@

Applying the terminal category preserving 2-functor $E \colon \mathbf{Cat} \to \mathbf{Cat}$ to this diagram yields:

@@

\section{From pseudofunctors to left adjoints}
In this section we investigate the necessary conditions to obtain a category $\mathrm{X}$ and a left adjoint $L \colon \mathrm{A} \to \mathrm{X}$ from a category $\mathrm{A}$ with finite colimits and a symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \mathbf{Cat}$. The main tool that allows us to do this will be the Grothendieck construction which we recall:

\begin{defn}
Let $\mathbf{Cat}_{\textnormal{lax},\star}$ denote the \define{2-category of lax-pointed categories} which has:
\begin{enumerate}
\item{objects as pairs $(\mathrm{A},c)$ where $\mathrm{A}$ is a category and $c$ is an object of $\mathrm{A}$, and}
\item{a morphism from $(\mathrm{A},c)$ to $(\mathrm{X},y)$ is a pair $(F,f)$ where $F \colon \mathrm{A} \to \mathrm{X}$ is a functor and $f \colon F(c) \to y$ is a morphism in $\mathrm{X}$.}
\end{enumerate}
\end{defn}

\begin{defn}
Given a pseudofunctor $F \colon \mathrm{A} \to \mathbf{Cat}$, the \define{Grothendieck construction of} $F$ is given by the `strict 2-pullback' of the following cospan:
 \[
\begin{tikzpicture}[scale=1.5]
\node (D) at (1,1) {$\int F$};
\node (A) at (0,0) {$\mathrm{A}$};
\node (B) at (1,-1) {$\mathbf{Cat}$};
\node (C) at (2,0) {$\mathbf{Cat}_{\textnormal{lax},\star}$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge[dashed] node[above]{$p$} (A)
(D) edge[dashed] node[above]{$q$} (C)
(A) edge node[below]{$F$} (B)
(C) edge node[below]{$P$} (B);
\end{tikzpicture}
\]
where $P \colon \mathbf{Cat}_{\textnormal{lax},\star} \to \mathbf{Cat}$ is the forgetful functor. This means that $\int F$ is a category which has:
\begin{enumerate}
\item{objects as pairs $(c, d \in F(c))$ and}
\item{a morphism from $(c, d \in F(c))$ to $(c^\prime, d^\prime \in F(c^\prime))$ is a pair $(f \colon c \to c^\prime,\alpha \colon F(f)(d) \to d^\prime)$.
